FMA: A Dataset For Music Analysis

### 2017-07-13 fma.tex
update reviewer 2

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset suitable for evaluating several tasks in MIR, a field concerned with browsing, searching, and organizing large music collections.
The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
The FMA aims to overcome this hurdle by providing \size GiB and \tduration days of Creative Commons-licensed audio from \ntracks tracks from \nartists artists and \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres.
It provides metadata like song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags; together with full-length, high-quality audio, and some pre-computed features. We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
Code, data, and usage examples are available at \url{\weblink}.
\end{abstract}

@@

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{\weblink}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most clips from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Volume is the number of samples times the dimensionality. Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

\begin{figure*}
	\begin{minipage}{0.74\linewidth}
		\centering
		\includegraphics[width=\linewidth]{growth.pdf}
	\end{minipage} \hfill
	\begin{minipage}{0.24\linewidth}
		\centering
		\includegraphics[width=\linewidth]{album_release_year.pdf}
	\end{minipage}
	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year, from 1902 to 2017.}
	\label{fig:growth}
	\label{fig:album_release_year}
\end{figure*}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{|llll|}
		\hline
		track\_id & title & number & \\
		information & language\_code & license & \\
		composer & publisher & lyricist & \\
		genres & genres\_all & genre\_top & \\
		duration & bit\_rate & interest & \\ %samplerate \\
		\#listens & \#comments & \#favorites & tags \\
		date\_created & date\_recorded & & \\
		\hline
		album\_id & title & type & \#tracks \\
		information & engineer & producer & \\
		\#listens & \#comments & \#favorites & tags \\
		date\_created & date\_released & & \\
		\hline
		artist\_id & name & & \\
		bio & members & \multicolumn{2}{l|}{associated\_labels} \\
		website & wikipedia\_page & \multicolumn{2}{l|}{related\_projects} \\
		location & longitude & latitude & \\
		& \#comments & \#favorites & tags \\
		date\_created & active\_year\_end & \multicolumn{2}{l|}{active\_year\_begin} \\
		\hline
	\end{tabular}
	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}.}
	%Not all of them are filled for each track.\footnote{See }
	%\texttt{albums.csv} and \texttt{artists.csv}.}
	% genres.csv ?
	\label{tab:metadata}
\end{table}

@@

% {\bf Available music datasets and their limitations.}
Unlike other modalities such as images or text for which a wealth of content is available, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
GTZAN \cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark MGR dataset. As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}. In addition to its small size, the main problems of GTZAN is mislabelling, repetitions, and distortions, as well as missing metadata (no artist name and song title) which prevent researchers to control for artists or album effects. % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} appear as contenders for a large-scale reference dataset. % They however each have limitations.
MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, audio features and raw audio. The poor audio quality and limited number of songs does however limit its usage.
MSD and AudioSet, while legal and very large, force researchers to download audio clips from online services.
% Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
% Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc. It has its own dedicated website with related code and additional datasets.
% 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals. 
On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.

@@

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
In comparison, FMA comes with legally redistributable full-length, high-quality (263 kbit/s on average) and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}
% Todo table quality

\textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}. All the metadata available through the API has been archived. In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
% which are necessary to evaluate content-based recommender systems.
There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}.
% social such as artist location or website
% images for tracks, albums and artists

\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio. No need to crawl the web and circumvent rate limits or access restrictions. Besides, we provide
% Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
some usage examples in the \texttt{usage.ipynb} Jupyter notebook to quick start using the data.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{duration_distribution.pdf}
	\caption{Track duration, from 0 to 3 hours (zoomed).}
	\label{fig:duration_distribution}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{listens_distribution.pdf}
	\caption{Album listens, from 0 to 3.6 millions (zoomed).}
	\label{fig:listens_distribution}
\end{figure}

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality mp3-encoded tracks, hand-picked by established audio curators. Each track is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio. Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}

@@

% Collection
As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks). See \figref{fig:growth} to visualize the growth of the dataset. In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}. Finally, mp3 audio was downloaded over HTTPS.
% Cleaning
Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71. Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.

@@

\begin{figure}
	\begin{minipage}[t]{0.44\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
	\end{minipage} \hfill
	\begin{minipage}[t]{0.55\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genres_per_track.pdf}
	\end{minipage}
	\caption{(left) Example of genre hierarchy for the top-level Soul-RnB genre. Left number is the \texttt{genre\_id}, right number is the number of tracks per genre. (right) Number of genres per track, from 0 to 31 (\texttt{genres\_all}).}
	\label{fig:genre_hierarchy}
	\label{fig:genres_per_track}
\end{figure}

@@

The collected track, album and artist meta-informations\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks. For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns. From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage. The audio for each track is stored in a file which name is the track id. All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.

%\todo{describe the metadata}

@@

While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by our source. Indeed, the website offers a parent genre for each of their 163 genres. \tabref{tab:genres} shows an excerpt of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree. \figref{fig:genre_hierarchy} shows an excerpt of the tree.
In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist. Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots. Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered in natural categories, it obeys a power law.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{tag_distribution.pdf}
	\caption{Per-track, album and artist tags, from 0 to 150.}
	\label{fig:tag_distribution}
\end{figure}

>>>>

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset suitable for evaluating several tasks in MIR, a field concerned with browsing, searching, and organizing large music collections.
The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
The FMA aims to overcome this hurdle by providing \size GiB and \tduration days of Creative Commons-licensed audio from \ntracks tracks from \nartists artists and \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres.
It provides full-length and high-quality audio, pre-computed features, together with track- and user-level metadata, tags, and free-form text such as biographies.
We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
Code, data, and usage examples are available at \url{\weblink}.
\end{abstract}

@@

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{\weblink}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most clips from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length [s] $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Volume is the number of samples times the dimensionality. Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

\begin{figure*}
	\begin{minipage}{0.74\linewidth}
		\centering
		\includegraphics[width=\linewidth]{growth.pdf}
	\end{minipage} \hfill
	\begin{minipage}{0.24\linewidth}
		\centering
		\includegraphics[width=\linewidth]{album_release_year.pdf}
	\end{minipage}
	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year (min 1902, max 2017).}
	\label{fig:growth}
	\label{fig:album_release_year}
\end{figure*}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{|r@{ }l@{\hspace{.8em}}r@{ }l@{\hspace{.8em}}r@{ }l|}
		\hline
		100\% & track\_id     & 100\% & title          &  93\% & number \\
		  2\% & information   &  14\% & language\_code & 100\% & license \\
		  4\% & composer      &   1\% & publisher      &   1\% & lyricist \\
		 98\% & genres        &  98\% & genres\_all    &  47\% & genre\_top \\
		100\% & duration      & 100\% & bit\_rate      & 100\% & interest \\ %samplerate \\
		100\% & \#listens     &   2\% & \#comments     &  61\% & \#favorites \\
		100\% & date\_created &   6\% & date\_recorded &  22\% & tags\\
		\hline
		100\% & album\_id     & 100\% & title          &     &  \\
		 94\% & type          &  96\% & \#tracks       &     &  \\
		 76\% & information   &  16\% & engineer       &  18\% & producer \\
		 97\% & \#listens     &  12\% & \#comments     &  38\% & \#favorites \\
		 97\% & date\_created &  64\% & date\_released &  18\% & tags \\
		\hline
		100\% & artist\_id    & 100\% & name           &  25\% & members \\
		 38\% & bio           &   5\% & \multicolumn{2}{@{}l}{associated\_labels} &  \\
		 43\% & website       &   2\% & wikipedia\_page & &  \\
		& & 5\% & \multicolumn{2}{@{}l}{related\_projects} & \\
		 37\% & location      &  23\% & longitude      &  23\% & latitude \\
		 11\% & \#comments    &  48\% & \#favorites    &  10\% & tags \\
		 99\% & date\_created &   8\% & active\_year\_begin & & \\
		& & 2\% & \multicolumn{2}{@{}l}{active\_year\_end} & \\
		\hline
	\end{tabular}
	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}. Percentages indicate coverage over all tracks, albums, and artists.}
	%Not all of them are filled for each track.\footnote{See }
	%\texttt{albums.csv} and \texttt{artists.csv}.}
	% genres.csv ?
	\label{tab:metadata}
\end{table}

@@

% {\bf Available music datasets and their limitations.}
Unlike other modalities such as images or text for which a wealth of content is available, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
GTZAN \cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark MGR dataset. As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}. In addition to its small size, the main problems of GTZAN is mislabelling, repetitions, and distortions, as well as missing metadata (no artist name and song title) which prevent researchers to control for artists or album effects. % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} and AcousticBrainz \cite{acousticbrainz} appear as contenders for a large-scale reference dataset. % They however each have limitations.
MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, audio features and raw audio. The poor audio quality and limited number of songs does however limit its usage.
MSD and AudioSet, while very large, force researchers to download audio clips from online services. AcousticBrainz's approach to the copyright issue is to ask the community to upload music descriptors of their tracks. Although it is the largest database to date, it will never distribute audio.
% Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
% Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc. It has its own dedicated website with related code and additional datasets.
% 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals. 
On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.

@@

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}. They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
In comparison, FMA comes with CC-licensed, full-length, high-quality and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}

\textbf{Metadata rich.} \todo{redundant with content} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}. All the metadata available through the API has been archived. In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
% which are necessary to evaluate content-based recommender systems.
There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}, while GTZAN offers none at all.
% social such as artist location or website
% images for tracks, albums and artists

\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio. No need to crawl the web and circumvent rate limits or access restrictions. Besides, we provide
% Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
some usage examples in the \texttt{usage.ipynb} Jupyter notebook to start using the data quickly.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{duration_distribution.pdf}
	\caption{Track duration (min 0, max 3 hours).}
	\label{fig:duration_distribution}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{listens_distribution.pdf}
	\caption{Album listens (min 0, max 3.6 millions).}
	\label{fig:listens_distribution}
\end{figure}

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
%The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality tracks, hand-picked by established audio curators. Each track is legally free to download as artists decided to release their works under a permissive license.
%Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}

@@

% Collection
As of April 1st 2017, when the dataset was gathered, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks). \figref{fig:growth} shows the growth of the dataset. In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.} Finally, mp3 audio was downloaded over HTTPS.
% Cleaning
Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71. Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.

While it may be argued that the dataset should be cleaned, we wanted it to resemble real world data. As such, we did not remove tracks which have too many genres, are too long, belong to rare genres, etc. Moreover, it is hard to set a threshold and the small number of outliers will not impact much the prediction accuracy. Researchers are obviously free to discard any track for training.

@@

\begin{figure}
	\begin{minipage}[t]{0.44\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
	\end{minipage} \hfill
	\begin{minipage}[t]{0.55\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genres_per_track.pdf}
	\end{minipage}
	\caption{(left) Example of genre hierarchy for the top-level Soul-RnB genre. Left number is the \texttt{genre\_id}, right number is the number of tracks per genre. (right) Number of genres per track.}
	\label{fig:genre_hierarchy}
	\label{fig:genres_per_track}
\end{figure}

@@

The metadata collected from the archive\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} was cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}. All the metadata has been produced by the artists when uploading their music. While the content is curated, the curators focus on the musical content not the metadata.
Metadata includes song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags. Coverage varies across fields, see \tabref{tab:metadata}.
For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns. From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage.

\todo{describe the metadata}
In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag. \figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name. We thus subtracted one.}

The audio for each track is stored in a file which name is the track id. All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kbit/s (263 kbit/s on average), and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
%todo: graph for audio quality

@@

\todo{The high coverage of genre information and fine-graine... makes the FMA well suited for MGR.}
Labeling noise is unavoidable
The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual tracks, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves. While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count. Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).

While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by the archive, which is the one the authors had in mind when annotating their tracks. That hierarchy is composed of \ngenres genres of which 16 are roots, the others being sub-genres. \tabref{tab:genres} shows an excerpt of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree. \figref{fig:genre_hierarchy} shows an excerpt of the tree.
\todo{3 genres limit}

In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist. Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots. Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre.
%As often encountered in natural categories, it obeys a power law.

While root genre recognition should be treated as a multi-label problem in general, \figref{fig:genres_per_track} shows that half the tracks have a single root genre.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{tag_distribution.pdf}
	\caption{Per-track, album and artist tags (min 0, max 150).}
	\label{fig:tag_distribution}
\end{figure}
